---
title: "SQUIRM Ratios"
author: "Grant Foster"
date: "2025-11-10"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

With the data from the GUTS system successfully parsed, you look forward to finally setting up your workstation's desktop with an approporiately helminth-themed background. However, before you can get started the cestode research division's lead engineer runs up to you in a panic. They let you know that the Soil Quantification and Identification Robot for Macroparasites (SQUIRM) unit has just broken down, and they need immediate help getting it back up and running!

The engineer explains that a part seems to be missing from the engine, but nobody can figure out which one. However, if you can add up all the part numbers in the engine schematic, it should be easy to work out which part is missing.

The engine schematic (your puzzle input) consists of a visual representation of the engine. There are lots of numbers and symbols you don't really understand, but apparently any number adjacent to a symbol, even diagonally, is a "part number" and should be included in your sum. (Periods (.) do not count as a symbol.)

Here is an example engine schematic:

467..114.. \
..#....... \
..35..633. \
......#... \
617$...... \
.....+.58. \
..592..... \
......755. \
...$.*.... \
.664.598.. \

In this schematic, two numbers are not part numbers because they are not adjacent to a symbol: 114 (top right) and 58 (middle right). Every other number is adjacent to a symbol and so is a part number; their sum is 4361.

Of course, the SQUIRM unit is a highly complex machine, and the engine schematic is much larger. What is the sum of all of the part numbers in the engine schematic?

```{r}
load("input.RDA")
```

Part II here? 

Adapted From: https://adventofcode.com/2023/day/3
